---
title: "Exercise: Explore, Clean, and Wrangle Data"
author: Kenji
format: html
---
## About the data

The data, "SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012," can be found on the EDI Data Portal[^1].

## Setup

```{r}
#| include: false
library(tidyverse)

```

## Read and explore data

```{r}
# download data
f1 <- "Time-series of lobster abundance and size.csv"
if(!(f1 %in% list.files("data"))){
  download.file(
    url = "https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.77.8&entityid=f32823fba432f58f66c06b589b7efac6",
    destfile = file.path("data", f1)
  )
}

f2 <- "Time-series of lobster trap buoy counts .csv"
if(!(f2 %in% list.files("data"))){
  download.file(
    url = "https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-sbc.77.8&entityid=66dd61c75bda17c23a3bce458c56ed84",
    destfile = file.path("data", f2)
  )
}

# read data in
lobster_abundance <- read_csv(file.path("data", f1))

lobster_traps <- read_csv(file.path("data", f2))
```
### Explore

```{r}
glimpse(lobster_abundance)
```

```{r}
glimpse(lobster_traps)
```


## Exercise 1 

The variable `SIZE_MM` uses `-99999` as the code for missing values (see metadata). This has the potential to cause conflicts with our analyses. Modify the data following these steps:

1. Verify the SIZE_MM variable contains -99999 values using unique().
2. Convert every -99999 value to an NA value using mutate() and na_if(). Look up the help page to see how to use na_if().
3. Check your output data using unique().

```{r}
# verify value
lobster_abundance %>%
  filter(SIZE_MM == -99999) %>%
  nrow()
```
```{r}
# modify na values
lobster_abundance <- lobster_abundance %>%
  # mutate(SIZE_MM = ifelse(SIZE_MM == -99999, as.numeric(NA), SIZE_MM)) %>%
  # alternative approach to above line
  mutate(SIZE_MM = na_if(SIZE_MM, -99999))
```

```{r}
# check
-99999 %in% unique(lobster_abundance$SIZE_MM)
```


## Exercise 2

Create a subset with the data for the lobsters at Arroyo Quemado (AQUE) that have a carapace length greater than 70 mm.

```{r}
sub <- lobster_abundance %>%
  filter(SITE == "AQUE", SIZE_MM > 70)

```

Create a subset with the traps’ information at all sites where abundance data is not NA. Note that you first have to identify which are these sites.

```{r}
sites <- lobster_abundance %>%
  pull(SITE) %>%
  unique()

traps2 <- lobster_traps %>%
  filter(SITE %in% sites)
```


## Exercise 3
## Exercise 4 
## Exercise 5
## Exercise 6

## References
[^1]: Santa Barbara Coastal LTER, D. Reed, and R. Miller. 2022. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/25aa371650a671bafad64dd25a39ee18 (Accessed 2023-06-29).